Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-05-31 paper_jasa.tex
Working on the lineup size and number of observers section

\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$, with an exact $x_{\alpha}$. For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$. It can be set to be smaller, even arbitrarily small, by increasing $K$. Type II error is harder to calculate, as is usually the case. In visual inference, individual abilities need to be incorporated to calculate Type II error. This is not necessary for Type I error, because it is simply the chance that the observer picks the actual data plot, by chance. For Type II error, we need to calculate the chance that the observer sees that actual data plot as different, when it really is different. This involves understanding the individual's visual skills. Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise, where $p_i$ is the probability that individual $i$ picks out the actual data plot from the lineup. If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}(K, p)$. Ee can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup. If there is evidence for individual skills influencing the probability $p_i$, then $X = \sum_i X_i$ will have a Poisson-Binomial distribution. This leads to the definition of the power of the visual test, as follows:

@@

%\red{How are $\Theta_0$ and $\Theta^c_0$ defined? - why do we need them? wouldn't it be better to re-formulate power in terms of acceptance region, not in terms of the parameter space? - $1/m$ also doesn't look right for the power - I think that's just looking at one observer.}
%\begin{dfn} \label{dfn:power}
%The \textbf{power} of a visual test $V_{\theta}$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
%    \begin{equation*}
%      \text{Power}_V(\theta)= 
%        \begin{cases} 
%              \frac1m & \text{if $\theta \in \Theta_0$,} \\
%              Pr(\text{Reject } H_0) &\text{if $\theta \in \Theta^c_0$.}
%        \end{cases}
%    \end{equation*}
%\end{dfn}
%
%
%\red{Alternative Power definition:}
\begin{dfn} \label{dfn:power}
The \textbf{power} of a visual test $V_{\theta}$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
    \begin{equation*}
      \text{Power}_V(\theta)= Pr(\text{Reject } H_0) = P(X \ge x_{1-\alpha}) = 
        \begin{cases} 
               1 - \text{Binom}_{K, 1/m}(x_{1-\alpha} - 1) & \text{ if } H_0 \text{ is true }.\\
              1 - F_{X, \theta} (x_{1-\alpha} - 1) & \text{ if } H_1 \text{ is true}.
        \end{cases}
    \end{equation*}
\end{dfn}

@@

\green{Done to here - the next parts are pretty long!}

@@

Let us make the assumption that an onlooker is able to identify  the chart corresponding to the data with the smallest $p$-value. Further we will assume that all onlookers  have the same ability in identifying the data plot.

@@

Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an onlooker will pick the data plot in a given lineup is 
\[
P(Z=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
\]
\end{proof}

@@

\blue{Note that the right hand side of equation (\ref{eqn:power_estimate}) is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers. While this allows us to compute this value independently of the parametric setting, it also means that onlookers might pick out the data plot for a reason not related to the value of $\theta$.

>>>>

\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$, with an exact $x_{\alpha}$. For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$. It can be set to be smaller, even arbitrarily small, by increasing $K$. Type II error is harder to calculate, as is usually the case. In visual inference, individual abilities need to be incorporated to calculate Type II error. This is not necessary for Type I error, because it is simply the chance that the observer picks the actual data plot, by chance. For Type II error, we need to calculate the chance that the observer sees that actual data plot as different, when it really is different. This involves understanding the individual's visual skills. Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise, where $p_i$ is the probability that individual $i$ picks out the actual data plot from the lineup. If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}(K, p)$. We can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup. If there is evidence for individual skills influencing the probability $p_i$, then $X = \sum_i X_i$ will have a Poisson-Binomial distribution. \green{WHY??? And how does this relate to the logistic model below?} This leads to the definition of the power of the visual test, as follows:

@@

%\red{How are $\Theta_0$ and $\Theta^c_0$ defined? - why do we need them? wouldn't it be better to re-formulate power in terms of acceptance region, not in terms of the parameter space? - $1/m$ also doesn't look right for the power - I think that's just looking at one observer.}
%\begin{dfn} \label{dfn:power}
%The \textbf{power} of a visual test $V_{\theta}$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
%    \begin{equation*}
%      \text{Power}_V(\theta)= 
%        \begin{cases} 
%              \frac1m & \text{if $\theta \in \Theta_0$,} \\
%              Pr(\text{Reject } H_0) &\text{if $\theta \in \Theta^c_0$.}
%        \end{cases}
%    \end{equation*}
%\end{dfn}
%
%
%\red{Alternative Power definition:}
\begin{dfn} \label{dfn:power}
The \textbf{power} of a visual test $V_{\theta}$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
    \begin{equation*}
      \text{Power}_V(\theta)= Pr(\text{Reject } H_0) = P(X \ge x_{\alpha}) = 
        \begin{cases} 
               1 - \text{Binom}_{K, 1/m}(x_{\alpha} - 1) & \text{ if } H_0 \text{ is true }.\\
              1 - F_{X, \theta} (x_{\alpha} - 1) & \text{ if } H_1 \text{ is true}.
        \end{cases}
    \end{equation*}
\end{dfn}

@@

\section{Effect of Lineup Size and Number of Observers}

\green{Notation changes $i$ to $j$, and $p$ to $\pi$. Can someone please decide which to use, and make the appropriate changes? How does the ``Let $X_{j\ell} \sim B_{1, \pi_{j\ell}}$...'' below relate to the ``let $X_i$ be a binary random variable...'' above? It looks like this is a repeat. One needs to go once notation is clarified.}

@@

Let us make the assumption that an observer is able to identify  the chart corresponding to the data with the smallest $p$-value. Further we will assume that all observers  have the same ability in identifying the data plot.

@@

Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an observer will pick the data plot in a given lineup is 
\[
P(Z=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
\]
\end{proof}

@@

\blue{Note that the right hand side of equation (\ref{eqn:power_estimate}) is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers. While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.
